A practical review on linear and nonlinear global approaches to flow instabilities

### 2018-11-30 99_Documentation/ARTICLE_STABFEM/ARTICLE_ASME_Revision.tex
Changes of Diogo in the references and small errors in the text in general.

%%% first author
\author{D. Fabre$^{(a)}$,  V. Citro$^{(b,a)}$,  D. Ferreira Sabino$^{(a)}$, P. Bonnefis$^{(a)}$,  F. Giannetti$^{(b)}$ \& M. Pigou$^(a)$ 
    \affiliation{
	$^{(a)}$ Institut de Mécanique des fluides de Toulouse (IMFT), University of Toulouse \\
	$^{(b)}$ Dipartimento do Ingegneria (DIIN), Universit\`a} di Salerno. 
}

@@

\paragraph{Navier-Stokes equations and weak form}

We start from the general problem of a flow field $[{\bf u},p]$ satisfying the incompressible Navier-Stokes equations on a domain $\Omega$,
\begin{eqnarray} \label{NSprimitive}
\partial_t {\bf u} = {\cal NS} ({\bf u},p)
\equiv - {\bf u} \cdot \nabla {\bf u} - \nabla p + \frac{2}{Re}  \nabla \cdot {\mathsf{D}}({\bf u}),  \\
\nabla \cdot {\bf u} = 0,
\end{eqnarray}
with suitable boundary conditions on the frontier $\partial \Omega$ of the domain.
Here $ {\mathsf{D}}({\bf u}) $ is the rate-of-strain tensor defined as
$$
 {\mathsf{D}}({\bf u}) = 1/2
\left( \nabla {\bf u} +  \nabla^T  {\bf u}\right),
$$ 
In the framework of finite element methods, we need to write the equation in weak form.
Prior to this we define a scalar product as follows, for either scalar or vectorial quantities 
$\left< \phi_1, \phi_2 \right> $:
$$
\left< \phi_1, \phi_2 \right> = \int_\Omega \overline{\phi}_1 \cdot \phi_2   \mbox{ d} \Omega,
$$
The weak form of the Navier-Stokes equations is readily defined by introducing test functions 
$[{\bf v},q]$ associated with the momentum and continuity equations, and integrating over the domain\footnote{In the simple presentation given here we have omitted the issue of boundary conditions. Details on way boundary conditions can be incorporated in the weak formulation through integration by parts can be found in appendix D.}
\be{NSweak}
\forall [{\bf v},q], \quad \partial_t \left< {\bf v}, {\bf u}\right> = \left< {\bf v} , {\cal NS} ({\bf u},p) \right> + \left< q, \nabla \cdot {\bf u}\right>.
\ee

@@

We look for a steady base-flow $[{\bf u}_b,p_b]$ satisfying the steady Navier-Stokes equations, i.e. 
%${\cal NS} ({\bf u}_b,p_b) = 0$.
${\cal NS} ({\bf u}_b,p_b) = 0$.
Suppose that we have a {\em guess}  for the base flow $[{\bf u}_b^g,p_b^g]$  which almost satisfies the equations.  We look for a better approximation under the form
\be{Newton1}
[{\bf u}_b,p_b]  = [{\bf u}_b^g,p_b^g] + [\delta {\bf u}_b, \delta p_b].
\ee
Injecting (\ref{Newton1}) into the weak form (\ref{NSweak}) of the Navier-Stokes equations and linearizing leads to  
${\cal NS}  ({\bf u}_b^g,p_b^g) +  {\cal LNS} _{{\bf u}_b^g}(\delta {\bf u}_b,\delta p_b)$, which can also be written in weak form as :
\bes{Newton2}
&\left< {\bf v}, {\cal NS} ({\bf u}_b^g,p_b^g)\right> + \left<q, \nabla \cdot {\bf u}_b^g\right>  
\\
+ &\left< { \bf v}, {\cal LNS} _{{\bf u}_b^g}( \delta {\bf u}_b,\delta p_b) \right> + \left<q, \nabla \cdot \delta{\bf u}_b\right> = 0,
\ees
where ${\cal LNS}$ is the linearised Navier-Stokes operator, defined by its action on a flow field $[{\bf u} , p]$ as follows 
\be{defNSL}
 {\cal LNS}_{{\bf U}}( {\bf u}, p) = - {\cal C}( {\bf U} , {\bf u}) -\nabla p
+\frac{2}{Re} \nabla  \cdot {\mathsf{D}}({\bf u}), %\quad ( \mbox{ with } \nabla \cdot {\bf u } = 0 \, ).
 \ee
and ${\cal C}$ is the convection operator defined by 
\be{defC}
{\cal C}( {\bf U} , {\bf u}) = \left( {\bf U} \cdot \nabla \right) {\bf u} + \left( {\bf u} \cdot \nabla \right)  {\bf U}.
\ee
This problem can now be discretized by projecting upon a basis of Taylor-Hood $(u,v,p) \rightarrow (P2,P2,P1)$ finite elements. Noting $\delta X$ the discretization of $[\delta {\bf u}_b,\delta p_b]$ this eventually leads to a matricial problem of the form $A \cdot \delta X = Y$. The procedure of Newton iteration is to solve iteratively this set of equations up to convergence.
In our implementation, the algorithm for a 2D incompressible flow is written in the Freefem++ solver {\sf Newton\_2D.edp} (an extract of this code featuring the implementation of the Newton loop is given in appendix E). This FreeFem++ solver is wrapped by the Octave/Matlab driver {\sf SF\_BaseFlow.m}.

@@

\paragraph{Adjoint eigenvalue problem, structural sensitivity, and wavemaker characterization}
Developed in the two past decades, the concept adjoint eigenmodes has now become an unavoidable complement to the linear global stability approach. We here give a short summary of the definition and usefulness of this concept and refer the reader to Luchini \& Bottaro\cite{LucBott2014} for further details.
First of all, the {\em adjoint linearised Navier-Stokes operator} ${\cal LNS}^\dag$ is defined thanks to the following 
property:
\bes{NSLAdj}
\forall ( {\bf u}, p ; {\bf v}, q), & \left< {\cal LNS}^\dag_{\bf U}( {\bf v},q) ,{\bf u}\right> + \left< \nabla \cdot {\bf v},p\right>  \\
=& \left< {\bf v}, {\cal LNS}_{\bf U} ({\bf u},p)\right> + \left< q, \nabla \cdot {\bf u}\right>.
\ees
We can then define the adjoint eigenvalues and eigenmodes as the solutions of the eigenvalue problem 
\be{EigenAdj} 
\forall ( {\bf u}, p), \quad  \lambda^\dag \left< \hat{\bf v}, {\bf u}\right> =
 \left< {\cal LNS}^\dag_{\bf U}( \hat{\bf v},\hat{q}) ,{\bf u}\right> + \left< \nabla \cdot \hat{\bf v},p\right> . 
\ee
It can be shown \cite{SchmidHenningson2001} that the adjoint eigenvalues $\lambda^\dag_k$ are the complex conjugates of the direct eigenvalues $\lambda_k$.

@@

To build an optimal mesh for the base-flow calculation, the idea is to use as the objective function $u_h$ the solution ${\bf u}_b$ itself, as computed on a previous mesh.%The procedure thus ensures that the mesh will be locally adapted to the gradients of the base flow. 
The base flow is then recomputed on the adapted mesh, providing a better approximation of the solution. The procedure can be repeated a few steps to ensure a right convergence.

@@

In the past decade, efforts have been devoted to extend the range of validity of global approaches into the nonlinear regime for $Re>Re_c$, with the double objective to describe the properties of the limit cycle reached after saturation and to derive amplitude equations describing the transitent dynamics towards this cycle. The two main milestones in this direction  are the {\em weakly nonlinear model}  (WNL) of Sipp \& Lebedev \cite{SippLebedev} and the {\em self-consistent model} (SC) of Manti\v{c}-Lugo et al \cite{MLugo2014}.
 In this review, we will address only the first of the two questions, namely  the description of the saturated cycle, and leave aside the question of transient dynamics. We will successively review the two aforementioned models, in a simplified formulation devoted to describe only the saturated cycle. We also provide a simple implementation of these two models. Finally, in line with the previous section, we will again illustrate with results obtained for the wake of a cylinder.

 In this part, to simplify the notations, we symbolically write the Navier-Stokes equations as $\partial_t {\bf u} = {\cal NS} ({\bf u})$, therefore dropping the systematic reference to the incompressibility constraint and associated pressure field. The same is done with the linearized operator ${\cal LNS}_{{\bf U}}( {\bf u})$.

@@

In the next sections we will document the predictions of the WNL and SC models regarding the quantities $A_E$ and $L_x$ adopted in past studies. In addition, we will document two other quantities of practical interest : the Drag and Lift forces $F_x$ and $F_y$ exerted on the cylinder. 
Both are periodic functions of time and; owing to symmetry consideration, the drag contains only even harmonics and the lift only odd harmonics:

@@

We first review the weakly nonlinear model of Sipp \& 
Lebedev\cite{SippLebedev}, also discussed by Gallaire et al. \cite{FDR2016}.  
The initial derivation of \cite{SippLebedev} makes use of a multiple scale method in order to obtained an amplitude equation.
This complete analysis is reproduced in appendix B. In the present paragraph, we give a simplified derivation of this model restricted to the description of the periodic saturated cycle.
The starting point can be taken as the following expansion of the velocity flow field:
\be{WNL1}
\begin{aligned}
&{\bf u} = {}  {\bf u}_{bc} + \epsilon \left[ A_{wnl}  \hat{\bf u} e^{i (\omega_c+\epsilon^2 \omega_\epsilon)  t} + c.c. \right]\\
&+ \epsilon^2 \left[ {\bf u}_\epsilon + |A_{wnl}|^2  {\bf u}_{2,0} + \left(  A_{wnl}^2 {\bf u}_{2,2} e^{2 i(\omega_c+\epsilon^2 \omega_\epsilon)  t} + c.c. \right) \right]\\
& + {\cal O}( \epsilon ^3),
\end{aligned}
\ee

@@

Injecting this ansatz into the Navier-Stokes equations and taking the mean value and the first Fourier component leads to the following coupled equations:
\begin{subequations}\label{eq_principal}
\begin{eqnarray}
{\cal NS}(  {\bf u }_m ) = \frac{{\cal C}( {\bf u }_{1,c}, {\bf u }_{1,c}) +{\cal C}( {\bf u }_{1,s}, {\bf u }_{1,s})}{4}, 
\label{HBEq0}
\\
 \omega {\bf u }_{1,s} =  {\cal LNS}_{{\bf u }_m}(  {\bf u }_{1,c} ),
\label{HBEq1}
\\
 -\omega {\bf u }_{1,c} =  {\cal LNS}_{{\bf u }_m}(  {\bf u }_{1,s} ).
\label{HBEq1b}
\end{eqnarray}
\end{subequations}

@@

Note that this system of equations contains three unknown fields (discretized on the Taylor-Hood basis) plus an extra scalar unknown, 
namely the frequency $\omega$. 
In order to solve this coupled problem, we thus need an extra scalar equation. The latter is provided by fixing {\em phase} of the cycle. Several choices are possible, but a convenient one is to decide that the instant $t=0$ corresponds to a maximum of the lift force.
This condition reads:
\be{HBphase}
{\cal L}_{Re}({\bf u}_{1s},p_{1s}) =0, 
\ee
so that the Fourier decomposition of the lift force \ref{drag_lift_def} will contain only a cosine term, namely 
\be{FySC}
F_y \approx F_{y,1,c} \cos \omega t, \quad \mbox{ with } \quad F_{y,1,c} = {\cal L}_{Re}({\bf u}_{1c},p_{1c}).
\ee

@@

When reformulated in this way, the nonlinear problem addressed by the self-consistent model actually reduces to the restriction to order one of a more general class of methods called  harmonic balance methods (HB) (DIOGO PUT A FEW REFERENCES HERE \cite{}). The development of such methods, or related ones such the time-spectral method (REFS ?)  and the search for efficient numerical implementation to solve such problems is currently an active field of development in the fluid dynamics community. When considering HB development with a large order of terms, direct resolution by inversion of the full matrix is not achievable. Methods including iterative solvers and efficient preconditioning are currently under development (O. Marquet, personal communication).

@@

After discretization, this leads to a linear problem of the form $A X = Y$ where $X$ is the discretized version of the unknowns  $[\delta {\bf u }_m, \delta {\bf u }_{1,c}, \delta {\bf u }_{1,s}, \delta \omega]$, and $A$ is a matrix of dimension $3 N_{dof} +1$, where $N_{dof}$ is the dimension of the Taylor-Hood basis of finite elements describing each of the three $[{\bf u},p]$ fields.  This system is solved iteratively in the same way as explained in section 2.1 for determination of the base flow. %Using the adapted mesh described in figure 2, the dimension of the system is $3 N_{dof} +1 \approx 58 000$ and the resolution remains possible at very reasonable computational costs. 
Mesh convergence issues for the nonlinear calculations are discussed in appendix A. It is shown that the mesh $\mathbf{M}_2$ designed using 'S' strategy is sufficient to yield converged results within $0.3\%$ accuracy for all quantities considered here, with the exception of the energy-amplitude $A_E$ which requires the more refined mesh $\mathbf{M}_4$ obtained using 'D' strategy.\footnote{In terms of computational cost, on a standard laptop the program  \ref{fig:listingNL} requires approximatively 413 seconds using  mesh $\mathbf{M}_2$ and 31 minutes using mesh $\mathbf{M}_4$.}

@@

Figure \ref{fig:MF60} illustrates the structure of the mean flow for $Re=60$. As already identified by  \cite{MLugo2014}, the recirculation region associated with this mean flow is notably shorter than the one associated with the base flow (figure 2).

%Line 19 shows how to plot the results along with the

%\ref{HBEq0,HBEq1,HBphase}

@@

{\em DIOGO : replace this paragraph by a description of figures displaying the comparison
The predictions of the SC models were also compared to numerical simulations results showing excellent agreement in the range $Re \in [Re_c,100]$. Regarding the forces exerted on the body, we can remark that the predictions of the WNL and HB1/SC models rapidly depart from each other as soon as $Re-Re_c \gtrsim 1$. Comparison of these results with DNS simulations is not provided here (consistently with the objectives of the present paper) but the reader may verify that the SC model correctly predicts these quantities in the range of Reynolds number considered here.
}

\cite{williamson1988defining}

@@

\begin{table*}
$$
\begin{array}{|c||c|c|c|c|c|c|c|c|c|}
\hline
\mbox{Mesh} & N_p & N_{dof} & \delta_{min} & \delta_{max} & \delta_A  & \delta_B  & \delta_C  & \delta_D  \\
\hline
\mathbf{M}_1 \mbox{(Adapt on base flow) } & 1429 & 12545	& 0.0131 & 14.33 		& 0.0259 	& 0.514 	& 0.819 	& 1.067 \\   
\mathbf{M}_2 \mbox{(Adapt on sensitivity) } & 2038 & 17885	& 0.0155 & 14.17 		& 0.02826 & 0.2046 	& 0.3909 	& 1.2014 	\\
\mathbf{M}_2' \mbox{(Adapt on sensitivity, split) } & 7974 & 70651	& 0.00786 & 7.131  		& 0.0104 	& 0.0744 	&   0.0975 & 0.614   \\
\mathbf{M}_3 \mbox{(Adapt on endogeneity) } & 3749 & 32655	& 0.01542 & 12.7109  		& 0.09689 	& 0.1236 	&   0.2004 & 0.65775   \\
\mathbf{M}_4 \mbox{(Adapt on mode) } 	& 12080  & 103564		&0.00825	& 12.461		& 0.0229	& 0.143	& 0.0993	& 0.0934 \\
\mathbf{M}_5 \mbox{(Adapt on adjoint) } 	& 3813 	& 330715		& 0.0108	& 13.95		& 0.0176 	& 0.114	& 0.127	& 1.177  \\
\hline
\end{array}
$$
\caption{Description of meshes used for validation of mesh adaptation strategy : number of vertices $N_p$ ; number of degrees of freedom of the P2-P2-P1 Taylor-hood basis $N_{dof}$ ; cell size (minimum and maximum value, and value at four characteristic point A,B,C,D as defined in the text). 
 }
\label{tab:conv1}
\end{table*}

Fx = 0.6447; Lx = 4.0736
\begin{table*}
$$
\begin{array}{|c||c|c||c||c|c||c|c|c|}
\hline
\mbox{Mesh} & L_x & F_x & \lambda & L_{x,HB1} & F_{x,HB1} & \omega_{HB1}  & F_{y,1,c} & A_{E} \\
\hline
\mathbf{M}_1 & 4.0733 & 0.64310	& 0.047056 + 0.74416i 		& 2.6073  & 0.69277 	&  0.84334 & 0.1298 & 1.8301  \\
\mathbf{M}_2 & 4.075 &.  0.64348  	& 0.046719 + 0.74489i 		& 2.6153  & 0.69246 	& 0.84351 & 0.12789 & 1.7134 \\ 
\mathbf{M}_2' & 4.0772 & 0.64391 	& 0.04684 +  0.74511i 		& 2.6141  & 0.69338 	& 0.84383 & 0.12799 & 2.048   \\  
\mathbf{M}_3 & 4.0749 & 0.64370 	& 0.04673 + 0.74538i		& 2.6140     & 0.69317	& 0.84408 & 0.12808 & 2.064   \\ 
\mathbf{M}_4 & 4.0748 & 0.64402 	& 0.04676  + 0.74502i 		& 2.6130 	& 0.69034 	& 0.84371 & 0.12809 & 2.5905 \\
\mathbf{M}_5 & 4.0736 & 0.64470	& 0.046782+0.74534i		& 2.6142 	& 0.69399		& 0.84390 & 0.12800 & 1.8234 \\
\hline
\end{array}
$$
\caption{Results for mesh adaptation strategy ($Re = 60$) : Base-flow characteristics $L_x$ and $F_x$, linear eigenvalue $\lambda$, 
Nonlinear self-consistent model characteristics  $\omega_{HB1}$, $F_{y,1,c}$ and $A_E$. All the results can be optained using the Octave/Matlab script
{\sf SCRIPT\_CYLINDER\_MESHCONVERGENCE.m}. }
\label{tab:conv2}
\end{table*}

>>>>

%%% first author
\author{D. Fabre$^{(a)}$,  V. Citro$^{(b,a)}$,  D. Ferreira Sabino$^{(a)}$, P. Bonnefis$^{(a)}$,  F. Giannetti$^{(b)}$ \& M. Pigou$^{(a)}$ 
    \affiliation{
	$^{(a)}$ Institut de M\'ecanique des fluides de Toulouse (IMFT), University of Toulouse \\
	$^{(b)}$ Dipartimento do Ingegneria (DIIN), Universit\`a} di Salerno. 
}

@@

\paragraph{Navier--Stokes equations and weak form}

We start from the general problem of a flow field $[{\bf u},p]$ satisfying the incompressible Navier--Stokes equations on a domain $\Omega$,
\begin{eqnarray} \label{NSprimitive}
\partial_t {\bf u} = {\cal NS} ({\bf u},p)
\equiv - {\bf u} \cdot \nabla {\bf u} - \nabla p + \frac{2}{Re}  \nabla \cdot {\mathsf{D}}({\bf u}),  \\
\nabla \cdot {\bf u} = 0,
\end{eqnarray}
with suitable boundary conditions on the frontier $\partial \Omega$ of the domain.
Here $ {\mathsf{D}}({\bf u}) $ is the rate-of-strain tensor defined as
$$
 {\mathsf{D}}({\bf u}) = 1/2
\left( \nabla {\bf u} +  \nabla^T  {\bf u}\right),
$$ 
In the framework of finite element methods, we need to write the equation in weak form.
Prior to this we define a scalar product as follows, for either scalar or vectorial quantities 
$\left< \phi_1, \phi_2 \right> $:
$$
\left< \phi_1, \phi_2 \right> = \int_\Omega \overline{\phi}_1 \cdot \phi_2   \mbox{ d} \Omega,
$$
The weak form of the Navier--Stokes equations is readily defined by introducing test functions 
$[{\bf v},q]$ associated with the momentum and continuity equations, and integrating over the domain\footnote{In the simple presentation given here we have omitted the issue of boundary conditions. Details on way boundary conditions can be incorporated in the weak formulation through integration by parts can be found in appendix D.}
\be{NSweak}
\forall [{\bf v},q], \quad \partial_t \left< {\bf v}, {\bf u}\right> = \left< {\bf v} , {\cal NS} ({\bf u},p) \right> + \left< q, \nabla \cdot {\bf u}\right>.
\ee

@@

We look for a steady base-flow $[{\bf u}_b,p_b]$ satisfying the steady Navier--Stokes equations, i.e. 
%${\cal NS} ({\bf u}_b,p_b) = 0$.
${\cal NS} ({\bf u}_b,p_b) = 0$.
Suppose that we have a {\em guess}  for the base flow $[{\bf u}_b^g,p_b^g]$  which almost satisfies the equations.  We look for a better approximation under the form
\be{Newton1}
[{\bf u}_b,p_b]  = [{\bf u}_b^g,p_b^g] + [\delta {\bf u}_b, \delta p_b].
\ee
Injecting (\ref{Newton1}) into the weak form (\ref{NSweak}) of the Navier--Stokes equations and linearizing leads to  
${\cal NS}  ({\bf u}_b^g,p_b^g) +  {\cal LNS} _{{\bf u}_b^g}(\delta {\bf u}_b,\delta p_b)$, which can also be written in weak form as :
\bes{Newton2}
&\left< {\bf v}, {\cal NS} ({\bf u}_b^g,p_b^g)\right> + \left<q, \nabla \cdot {\bf u}_b^g\right>  
\\
+ &\left< { \bf v}, {\cal LNS} _{{\bf u}_b^g}( \delta {\bf u}_b,\delta p_b) \right> + \left<q, \nabla \cdot \delta{\bf u}_b\right> = 0,
\ees
where ${\cal LNS}$ is the linearised Navier--Stokes operator, defined by its action on a flow field $[{\bf u} , p]$ as follows 
\be{defNSL}
 {\cal LNS}_{{\bf U}}( {\bf u}, p) = - {\cal C}( {\bf U} , {\bf u}) -\nabla p
+\frac{2}{Re} \nabla  \cdot {\mathsf{D}}({\bf u}), %\quad ( \mbox{ with } \nabla \cdot {\bf u } = 0 \, ).
 \ee
and ${\cal C}$ is the convection operator defined by 
\be{defC}
{\cal C}( {\bf U} , {\bf u}) = \left( {\bf U} \cdot \nabla \right) {\bf u} + \left( {\bf u} \cdot \nabla \right)  {\bf U}.
\ee
This problem can now be discretized by projecting upon a basis of Taylor--Hood $(u,v,p) \rightarrow (P2,P2,P1)$ finite elements. Noting $\delta X$ the discretization of $[\delta {\bf u}_b,\delta p_b]$ this eventually leads to a matricial problem of the form $A \cdot \delta X = Y$. The procedure of Newton iteration is to solve iteratively this set of equations up to convergence.
In our implementation, the algorithm for a 2D incompressible flow is written in the Freefem++ solver {\sf Newton\_2D.edp} (an extract of this code featuring the implementation of the Newton loop is given in appendix E). This FreeFem++ solver is wrapped by the Octave/Matlab driver {\sf SF\_BaseFlow.m}.

@@

\paragraph{Adjoint eigenvalue problem, structural sensitivity, and wavemaker characterization}
Developed in the two past decades, the concept adjoint eigenmodes has now become an unavoidable complement to the linear global stability approach. We here give a short summary of the definition and usefulness of this concept and refer the reader to Luchini \& Bottaro\cite{LucBott2014} for further details.
First of all, the {\em adjoint linearised Navier--Stokes operator} ${\cal LNS}^\dag$ is defined thanks to the following 
property:
\bes{NSLAdj}
\forall ( {\bf u}, p ; {\bf v}, q), & \left< {\cal LNS}^\dag_{\bf U}( {\bf v},q) ,{\bf u}\right> + \left< \nabla \cdot {\bf v},p\right>  \\
=& \left< {\bf v}, {\cal LNS}_{\bf U} ({\bf u},p)\right> + \left< q, \nabla \cdot {\bf u}\right>.
\ees
We can then define the adjoint eigenvalues and eigenmodes as the solutions of the eigenvalue problem 
\be{EigenAdj} 
\forall ( {\bf u}, p), \quad  \lambda^\dag \left< \hat{\bf v}, {\bf u}\right> =
 \left< {\cal LNS}^\dag_{\bf U}( \hat{\bf v},\hat{q}) ,{\bf u}\right> + \left< \nabla \cdot \hat{\bf v},p\right> . 
\ee
It can be shown \cite{SchmidHenningson2001} that the adjoint eigenvalues $\lambda^\dag_k$ are the complex conjugates of the direct eigenvalues $\lambda_k$.

@@

To build an optimal mesh for the base-flow calculation, the idea is to use as the objective function $u_h$ the solution ${\bf u}_b$ itself, as computed on a previous mesh. %The procedure thus ensures that the mesh will be locally adapted to the gradients of the base flow. 
The base flow is then recomputed on the adapted mesh, providing a better approximation of the solution. The procedure can be repeated a few steps to ensure a right convergence.

@@

In the past decade, efforts have been devoted to extend the range of validity of global approaches into the nonlinear regime for $Re>Re_c$, with the double objective to describe the properties of the limit cycle reached after saturation and to derive amplitude equations describing the transient dynamics towards this cycle. The two main milestones in this direction  are the {\em weakly nonlinear model}  (WNL) of Sipp \& Lebedev \cite{SippLebedev} and the {\em self-consistent model} (SC) of Manti\v{c}-Lugo et al \cite{MLugo2014}.
 In this review, we will address only the first of the two questions, namely  the description of the saturated cycle, and leave aside the question of transient dynamics. We will successively review the two aforementioned models, in a simplified formulation devoted to describe only the saturated cycle. We also provide a simple implementation of these two models. Finally, in line with the previous section, we will again illustrate with results obtained for the wake of a cylinder.

 In this part, to simplify the notations, we symbolically write the Navier--Stokes equations as $\partial_t {\bf u} = {\cal NS} ({\bf u})$, therefore dropping the systematic reference to the incompressibility constraint and associated pressure field. The same is done with the linearized operator ${\cal LNS}_{{\bf U}}( {\bf u})$.

@@

In the next sections we will document the predictions of the WNL and SC models regarding the quantities $A_E$ and $L_x$ adopted in past studies. In addition, we will document two other quantities of practical interest : the Drag and Lift forces $F_x$ and $F_y$ exerted on the cylinder. 
Both are periodic functions of time and, owing to symmetry consideration, the drag contains only even harmonics and the lift only odd harmonics:

@@

We first review the weakly nonlinear model of Sipp \& 
Lebedev\cite{SippLebedev}, also discussed by Gallaire et al. \cite{FDR2016}.  
The initial derivation of \cite{SippLebedev} makes use of a multiple scale method in order to obtain an amplitude equation.
This complete analysis is reproduced in appendix B. In the present paragraph, we give a simplified derivation of this model restricted to the description of the periodic saturated cycle.
The starting point can be taken as the following expansion of the velocity flow field:
\be{WNL1}
\begin{aligned}
&{\bf u} = {}  {\bf u}_{bc} + \epsilon \left[ A_{wnl}  \hat{\bf u} e^{i (\omega_c+\epsilon^2 \omega_\epsilon)  t} + c.c. \right]\\
&+ \epsilon^2 \left[ {\bf u}_\epsilon + |A_{wnl}|^2  {\bf u}_{2,0} + \left(  A_{wnl}^2 {\bf u}_{2,2} e^{2 i(\omega_c+\epsilon^2 \omega_\epsilon)  t} + c.c. \right) \right]\\
& + {\cal O}( \epsilon ^3),
\end{aligned}
\ee

@@

Injecting this ansatz into the Navier--Stokes equations and taking the mean value and the first Fourier component leads to the following coupled equations:
\begin{subequations}\label{eq_principal}
\begin{eqnarray}
{\cal NS}(  {\bf u }_m ) = \frac{{\cal C}( {\bf u }_{1,c}, {\bf u }_{1,c}) +{\cal C}( {\bf u }_{1,s}, {\bf u }_{1,s})}{4}, 
\label{HBEq0}
\\
 \omega {\bf u }_{1,s} =  {\cal LNS}_{{\bf u }_m}(  {\bf u }_{1,c} ),
\label{HBEq1}
\\
 -\omega {\bf u }_{1,c} =  {\cal LNS}_{{\bf u }_m}(  {\bf u }_{1,s} ).
\label{HBEq1b}
\end{eqnarray}
\end{subequations}

@@

Note that this system of equations contains three unknown fields (discretized on the Taylor--Hood basis) plus an extra scalar unknown, 
namely the frequency $\omega$. 
In order to solve this coupled problem, we thus need an extra scalar equation. The latter is provided by fixing {\em phase} of the cycle. Several choices are possible, but a convenient one is to decide that the instant $t=0$ corresponds to a maximum of the lift force.
This condition reads:
\be{HBphase}
{\cal L}_{Re}({\bf u}_{1s},p_{1s}) =0, 
\ee
so that the Fourier decomposition of the lift force \ref{drag_lift_def} will contain only a cosine term, namely 
\be{FySC}
F_y \approx F_{y,1,c} \cos \omega t, \quad \mbox{ with } \quad F_{y,1,c} = {\cal L}_{Re}({\bf u}_{1c},p_{1c}).
\ee

@@

When reformulated in this way, the nonlinear problem addressed by the self-consistent model actually reduces to the restriction to order one of a more general class of methods called harmonic balance methods (HB). One of the first aplications of these methods in the domain of fluid mechanics has done by Carte and Du\v{s}ek (\cite{CarteDusek1995},\cite{CarteDusekfraunie1996}). The development of such methods, or related ones such the time-spectral method (\cite{HallThomasClark2002}) and the search for efficient numerical implementation to solve such problems is currently an active field of development in the fluid dynamics community. When considering HB development with a large order of terms, direct resolution by inversion of the full matrix is not achievable. Methods including iterative solvers and efficient preconditioning are currently under development (O. Marquet, personal communication).

@@

After discretization, this leads to a linear problem of the form $A X = Y$ where $X$ is the discretized version of the unknowns  $[\delta {\bf u }_m, \delta {\bf u }_{1,c}, \delta {\bf u }_{1,s}, \delta \omega]$, and $A$ is a matrix of dimension $3 N_{dof} +1$, where $N_{dof}$ is the dimension of the Taylor--Hood basis of finite elements describing each of the three $[{\bf u},p]$ fields.  This system is solved iteratively in the same way as explained in section 2.1 for determination of the base flow. %Using the adapted mesh described in figure 2, the dimension of the system is $3 N_{dof} +1 \approx 58 000$ and the resolution remains possible at very reasonable computational costs. 
Mesh convergence issues for the nonlinear calculations are discussed in appendix A. It is shown that the mesh $\mathbf{M}_2$ designed using 'S' strategy is sufficient to yield converged results within $0.3\%$ accuracy for all quantities considered here, with the exception of the energy-amplitude $A_E$ which requires the more refined mesh $\mathbf{M}_4$ obtained using 'D' strategy.\footnote{In terms of computational cost, on a standard laptop the program  \ref{fig:listingNL} requires approximatively 413 seconds using  mesh $\mathbf{M}_2$ and 31 minutes using mesh $\mathbf{M}_4$.}

@@

Figure \ref{fig:MF60} illustrates the structure of the mean flow for $Re=60$. As already identified by  \cite{MLugo2014}, the recirculation region associated with this mean flow is notably shorter than the one associated with the base flow (figure \ref{fig:Baseflow}).

%Line 19 shows how to plot the results along with the 
%\ref{HBEq0,HBEq1,HBphase}

@@

%replace this paragraph by a description of figures displaying the comparison

The present results are also compared with the available literature, namely with the experiments of \cite{williamson1988defining} for the $St$ number and  with the DNS results of \cite{MLugo2014} for $A_E$ and $L_x$.
The predictions of the SC/HB1 model show an excellent agreement in the range $Re \in [Re_c,100]$ with the rReferences. %Regarding the forces exerted on the body, 
However, we can remark that the predictions of the WNL model rapidly depart from the previous ones as soon as $Re-Re_c \gtrsim 1$.

%Comparison of these results with DNS simulations is not provided here (consistently with the objectives of the present paper) but the reader may verify that the SC model correctly predicts these quantities in the range of Reynolds number considered here.

@@

\begin{table*}
$$
\begin{array}{|c||c|c|c|c|c|c|c|c|c|}
\hline
\mbox{Mesh} & N_p & N_{dof} & \delta_{min} & \delta_{max} & \delta_A  & \delta_B  & \delta_C  & \delta_D  \\
\hline
\mathbf{M}_1 \mbox{(Adapt on base flow) } & 1429 & 12545	& 0.0131 & 14.33 		& 0.0259 	& 0.514 	& 0.819 	& 1.067 \\   
\mathbf{M}_2 \mbox{(Adapt on sensitivity) } & 2038 & 17885	& 0.0155 & 14.17 		& 0.02826 & 0.2046 	& 0.3909 	& 1.2014 	\\
\mathbf{M}_2' \mbox{(Adapt on sensitivity, split) } & 7974 & 70651	& 0.00786 & 7.131  		& 0.0104 	& 0.0744 	&   0.0975 & 0.614   \\
\mathbf{M}_3 \mbox{(Adapt on endogeneity) } & 3749 & 32655	& 0.01542 & 12.7109  		& 0.09689 	& 0.1236 	&   0.2004 & 0.65775   \\
\mathbf{M}_4 \mbox{(Adapt on mode) } 	& 12080  & 103564		&0.00825	& 12.461		& 0.0229	& 0.143	& 0.0993	& 0.0934 \\
\mathbf{M}_5 \mbox{(Adapt on adjoint) } 	& 3813 	& 330715		& 0.0108	& 13.95		& 0.0176 	& 0.114	& 0.127	& 1.177  \\
\hline
\end{array}
$$
\caption{Description of meshes used for validation of mesh adaptation strategy : number of vertices $N_p$ ; number of degrees of freedom of the P2-P2-P1 Taylor--Hood basis $N_{dof}$ ; cell size (minimum and maximum value, and value at four characteristic point A,B,C,D as defined in the text). 
 }
\label{tab:conv1}
\end{table*}

%Fx = 0.6447; Lx = 4.0736
\begin{table*}
$$
\begin{array}{|c||c|c||c||c|c||c|c|c|}
\hline
\mbox{Mesh} & L_x & F_x & \lambda & L_{x,HB1} & F_{x,HB1} & \omega_{HB1}  & F_{y,1,c} & A_{E} \\
\hline
\mathbf{M}_1 & 4.0733 & 0.64310	& 0.047056 + 0.74416i 		& 2.6073  & 0.69277 	&  0.84334 & 0.1298 & 1.8301  \\
\mathbf{M}_2 & 4.075 &.  0.64348  	& 0.046719 + 0.74489i 		& 2.6153  & 0.69246 	& 0.84351 & 0.12789 & 1.7134 \\ 
\mathbf{M}_2' & 4.0772 & 0.64391 	& 0.04684 +  0.74511i 		& 2.6141  & 0.69338 	& 0.84383 & 0.12799 & 2.048   \\  
\mathbf{M}_3 & 4.0749 & 0.64370 	& 0.04673 + 0.74538i		& 2.6140     & 0.69317	& 0.84408 & 0.12808 & 2.064   \\ 
\mathbf{M}_4 & 4.0748 & 0.64402 	& 0.04676  + 0.74502i 		& 2.6130 	& 0.69034 	& 0.84371 & 0.12809 & 2.5905 \\
\mathbf{M}_5 & 4.0736 & 0.64470	& 0.046782+0.74534i		& 2.6142 	& 0.69399		& 0.84390 & 0.12800 & 1.8234 \\
\hline
\end{array}
$$
\caption{Results for mesh adaptation strategy ($Re = 60$) : Base-flow characteristics $L_x$ and $F_x$, linear eigenvalue $\lambda$, 
Nonlinear self-consistent model characteristics  $\omega_{HB1}$, $F_{y,1,c}$ and $A_E$. All the results can be optained using the Octave/Matlab script
{\sf SCRIPT\_CYLINDER\_MESHCONVERGENCE.m}. }
\label{tab:conv2}
\end{table*}
